---
title: "Assignment 5"
author: "Ke'Shondra Reed"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(HelpersMG)
library(lubridate)
library(ggplot2)
library(kableExtra)
```

```{r harvData, echo = FALSE, message = FALSE}
if(!file.exists("./raw_data/Countries-Deaths.tsv")) {
	wget(url = "https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/L20LOT/FZLQRQ", destfile = "./raw_data/Countries-Deaths.tsv")
}
```


```{r countryName, echo = FALSE, message = FALSE}
dateButchery <- function(date) {
	return(mdy(paste(substr(date, 1, 2), "-", substr(date, 3, 4), "-2020")))
}

read_tsv("./raw_data/Countries-Deaths.tsv") %>%
	filter(CNTRYNAME == "Germany" | CNTRYNAME == "Italy" | CNTRYNAME == "France") %>%
	pivot_longer(-CNTRYNAME,
				 names_to = "Date",
				 values_to = "Total_Deaths") %>%
	rename(Country = CNTRYNAME) %>%
	mutate(Date = dateButchery(Date)) %>%
	write.table(file = "./output/GFI_total_deaths_by_date.csv", row.names = FALSE, sep = ",")
```


```{r kmrSnow, echo = FALSE, message = FALSE}
toMuch <- function(tibbleColumn) {
	count = 1
	for(element in tibbleColumn) {
		if(element < 0) {
			tibbleColumn[[count]] = 0
		}
		count = count + 1
	}
	return(tibbleColumn)
}

read_csv("./output/GFI_total_deaths_by_date.csv") %>%
	transmute(Country, Date, Daily_Deaths = toMuch(Total_Deaths - lag(Total_Deaths, default = 0))) %>%
	write.table(file = "./output/GFI_daily_deaths_by_date.csv", row.names = FALSE, sep = ",")
```




## Introduction 

The data from Harvard University showed a total of daily death rates from COVID-19. Beginning in February until June in Germany, France, and Italy. In late Feb. early March the number of total deaths had a steep rise in Italy. Whereas Germany and France began to rise gradual in a short time after.

```{r totalDeath, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(read_csv("./output/GFI_total_deaths_by_date.csv"), aes(x = Date, y = Total_Deaths, color = Country)) + geom_line(size = 1.5) + labs(y = "Total COVID-19 Deaths") + theme_linedraw() + scale_y_log10()
```

```{r dailyDeaths, echo = FALSE, message = FALSE}
ggplot(read_csv("./output/GFI_daily_deaths_by_date.csv"), aes(x = Date, y = Daily_Deaths, color = Country)) + geom_line(size = 1.5) + labs(y = "Daily COVID-19 Deaths") + theme_linedraw() + ylim(0,1000)
```

```{r bestInShow, echo = FALSE, message = FALSE}
read_csv("./output/GFI_daily_deaths_by_date.csv") %>%
	group_by(Country) %>%
	transmute(Max_Daily_Deaths = max(Daily_Deaths)) %>%
	distinct() %>%
	knitr::kable(col.names = c("Country", "Maximum Deaths per Day")) %>%
	kable_styling(full_width = FALSE, position = "left")
```


## Conlusion
Looking at the graph Italy deaths began before Germany and France, and the death rised very fast.Germany total death was lower than Italy and France. Even though France began a little after Italy the total death peaked the highest. In the beginning and late April the total death began to come decrease.


## Reference

China Data Lab, 2020, "Countries-Deaths.tab", World COVID-19 Daily Cases with Basemap, https://doi.org/10.7910/DVN/L20LOT/FZLQRQ, Harvard Dataverse, V22, UNF:6:AntG1aQ3uPjWCuJZ5P2uEA== [fileUNF]







